---
title: "data-viz"
author: "Joe DeCesaro"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#General use
library(tidyverse)
library(here)
library(janitor)

#Specific for data visualization
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)

library(sf)
```

## Read in the lizards data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

Histogram (really bad), of lizard weights

```{r}
ggplot(data = lizards, aes(x = weight)) +
         geom_histogram(fill = "orange", 
                        color = "purple",
                        size = 0.2,
                        linetype = "dotted")

# can also do it this way
ggplot() +
         geom_histogram(data = lizards, aes(x = weight))

# color is for points and lines, fill is for polygons
```
Lets make a horrible scatter plot
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(fill = "yellow", 
             shape = 25,
             color = "blue",
             alpha = 0.9)
```

Make a bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length
(This is never a good plot FYI)
```{r}
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name, 
                 size = total_length),
             shape = 22,
             fill = "black") +
             theme_minimal()
```

Make something not so terrible
```{r}
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name)

# Facet grid does something different
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name)) + 
  facet_grid(sex ~ tail)
```

Find lizards counts by common name
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

lizard_counts_cn_tail <- lizards %>% 
  count(common_name, tail)

ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) + 
  geom_col(aes(fill = commmon_name), show.legend = FALSE)
  
```

Try converting common name to an ordered factor outside of ggplot
```{r}
common_name_fac <- lizard_counts %>% 
  mutate(factor_names = fct_reorder(common_name, count))
```

## Axis scale breaks, limits and labels

Scatter plot: total_length (x), weight (y)
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) + 
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

Create a new version of lizards that counts observations by date

Transform the date column to class Date, then find counts of observations by date.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())

# Make a line plot (geom_line()) of date (x) and count (y)
ggplot(lizard_counts, aes( x = date, y = count)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")
```

