---
title: "data-viz"
author: "Joe DeCesaro"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#General use
library(tidyverse)
library(here)
library(janitor)

#Specific for data visualization
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)

library(sf)
```

## Read in the lizards data
```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

Histogram (really bad), of lizard weights

```{r}
ggplot(data = lizards, aes(x = weight)) +
         geom_histogram(fill = "orange", 
                        color = "purple",
                        size = 0.2,
                        linetype = "dotted")

# can also do it this way
ggplot() +
         geom_histogram(data = lizards, aes(x = weight))

# color is for points and lines, fill is for polygons
```
Lets make a horrible scatter plot
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(fill = "yellow", 
             shape = 25,
             color = "blue",
             alpha = 0.9)
```

Make a bubble plot where the color of the points changes based on common_name and the size of points varies based on total_length
(This is never a good plot FYI)
```{r}
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name, 
                 size = total_length),
             shape = 22,
             fill = "black") +
             theme_minimal()
```

Make something not so terrible
```{r}
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name)

# Facet grid does something different
ggplot(data = lizards, aes( total_length, y = weight)) + 
  geom_point(aes(color = common_name)) + 
  facet_grid(sex ~ tail)
```

Find lizards counts by common name
```{r}
lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

lizard_counts_cn_tail <- lizards %>% 
  count(common_name, tail)

ggplot(data = lizard_counts, aes(y = fct_reorder(common_name, count), x = count)) + 
  geom_col(aes(fill = commmon_name), show.legend = FALSE)
  
```

Try converting common name to an ordered factor outside of ggplot
```{r}
common_name_fac <- lizard_counts %>% 
  mutate(factor_names = fct_reorder(common_name, count))
```

## Axis scale breaks, limits and labels

Scatter plot: total_length (x), weight (y)
```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) + 
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```

Create a new version of lizards that counts observations by date

Transform the date column to class Date, then find counts of observations by date.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  group_by(date) %>% 
  summarize(count = n())

# Make a line plot (geom_line()) of date (x) and count (y)
ggplot(lizard_counts, aes( x = date, y = count)) +
  geom_line() +
  scale_x_date(date_breaks = "3 years",
               date_labels = "%Y")
```

## Adding a color gradient

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  scale_color_gradientn(colors = c("orange", "red", "purple", "black"))
# the above does not show anything at specific steps which is not ehlpful usually

ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) + 
  scale_color_stepsn(colors = c("green", "blue", "purple"), 
                     breaks = c(0, 20, 40, 50))
```

### Update color scheme using a palette in paletteer

Make a horizontal boxplot with common_name on the y-axis, total_length on the x-axis, with color changing based on common_name
```{r}
lizards_fct <- lizards %>% 
  mutate(common_name_fct = fct_reorder(common_name, .x = total_length, .fun = median))

ggplot(lizards_fct, aes(x = total_length, y = common_name_fct)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "colorBlindness::Blue2DarkOrange18Steps")
  

```

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  theme(
    plot.background = element_rect(fill = "cyan4"),
    panel.grid = element_blank(), 
    axis.text.x = element_text(color = "orange"),
    axis.title = element_text(color = "yellow", size = 15)
  )
```

## Day 9! - More ggplot stuff, ggrepel

```{r}
ww_lizards <- lizards %>% 
  filter(common_name == "western whiptail", 
         site == "sand")
```

```{r}
ggplot(ww_lizards, aes(x = total_length, y = weight)) +
  geom_point() + 
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)
```

View gapminder, filter for Europe creat a scatterplot of gdpPerCap (x) and lifeExp (y), labeled by country name
```{r}

gapminder %>% 
  filter(continent == "Europe", year == 2002) %>% 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3, max.overlaps = 20)

```

## gghighlight

```{r}
p <- ggplot(lizards, aes( x = total_length, y = weight)) +
  geom_point()

p + 
  gghighlight(toe_num == 250, label_key = toe_num)

```

```{r}
q <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_line(aes(color = common_name)) +
  gghighlight(max(weight) > 30)

q
```

## patchwork for compound figures
```{r}
(p | q) / q &
  theme_minimal()
```

## a few new graph types

### Marginal pltos

```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight)

ggplot(data = whiptails, aes(x = total_length, y = weight)) + 
  geom_point() + 
  geom_rug()
  

```

Marginal plot with boxplot
```{r}
my_plot <- ggplot(data = whiptails, aes(x = total_length, y = weight)) + 
  geom_point(aes(color = sex), size = 2) +
  scale_color_manual(values = c("cyan4", "black", "goldenrod"), 
                     name = "Sex", 
                     labels = c("Female", "Juvenile", "Male")) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggMarginal(my_plot, type = "boxplot", groupColour = TRUE)
```

## Beeswarm plot

```{r}
ggplot(data = whiptails, aes( x = sex, y = weight)) +
  geom_beeswarm() +
  geom_boxplot(fill = NA)
```

## A heatmap with geom_tile()

Make a new df named lizards_counts starting with lizards with data converted to Class Date. Then count by year and common name the number of lizards observed.
```{r}
lizards_count <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  mutate(year = lubridate::year(date)) %>% 
  group_by(year, common_name) %>% 
  summarize(count = n())
```

```{r}
ggplot(data = lizards_count, aes(x = year, y = common_name)) +
  geom_tile(aes(fill = count), na.rm = TRUE) +
  geom_text(aes(label = count), color = "white", size = 3) +
  scale_fill_viridis_c()
  # scale_fill_gradientn(colors = c("navy", "red", "orange"))
```

## Making a map!

Use `read_sf` to read in the doc.kml file.
```{r}
jornada_vegetation <- sf::read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  select(Name) %>% 
  clean_names()

ggplot(data = jornada_vegetation) +
  geom_sf(aes(fill = name), color = "NA") + 
  scale_fill_paletteer_d(palette = "ggthemes::manyeys") + 
  labs( x = "longitude",
        y = "Latitude", 
        fill = "Dominant Vegetation") + 
  theme_minimal()


```



